---
title: "A novel way to visulaize event time data"
date: "`r Sys.Date()`"
output: 
  html_document2:
    fig_caption: yes
author:  
  - name: Abdullah Al Mahmud
---

##### Continue on article repo

# Abstract

  Visualization is immensely important in statistical analysis and prediction. It helps in describing the data as well as in choosing the appropriate method of modeling and prediction. So far, however, there is no way to visualize event time data on a clock. In this paper we present a few novel ways to visualize various event time data on chart we call 'Data Clock'. Given events ocurring at various times in a 24-hour day, we show the events on a clock corresponding to the time of event. This makes it more useful to perceive exactly when the events occurred. In addition, we can easily compare the time of the event with other events. Time of events plotted on a clock chart is subtantially more revealing than a bar chart or a pie chart. It can demonstrate around which time events are less or more frequent as well the gap between the events. We also introduce __these functions__.


# Methodology

Fetch from package "clockchart-construction.Rmd" file. 

## Clock Skeleton

## Plotting Times on the Clock

# Examples

Let us explore US Accidents data. Although the main focus of these data 
seem to be on accident times, we cna employ it to visualize the comparison 
between the existing line chart and our proposed clock chart. 

```{r}
library(clockplot)
library(ggplot2)
acdt <- read.csv("https://raw.githubusercontent.com/mahmudstat/open-analysis/main/data/usacc.csv")
acdt <- acdt[,c(9:12)]
head(acdt)
```

Let us plot temperature against time.

```{r, echo=FALSE}
p <- ggplot2::ggplot(acdt, ggplot2::aes(Time, Temperature.F.))+
  ggplot2::geom_col(width = 0.2)+
  labs(title = "A traditional Line Chart Showing Accidents by Time", 
       y = "Temperature (°F)")
p + ggplot2::theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

We have no idea which temperature belongs to which time, although we have 
changed X- axis labels at right angles (i.e., vertically). 

Now let us take a look into a clock plot. 

```{r}
clock_chart(data = acdt, time = Time)
```

Although this is the simplest kind of clock chart that we propose, this still
makes a lot more sense. We clearly see a huge gap between around 9 PM to 
around 4:30 AM, during which period no reading of temperature was made (in
other example, this would mean no event occurred, which is actually relevant to
this data set as well, which records times of accidents. Why no accidents 
occurred during this period is subject to further discussion; perhaps, 
the traffic was less during this time.)

We can easily observe that most incidents occurred from 5 AM to 9 AM. Between 
other hours, roughly two accidents occurred. 


Let us try plotting 20 values only.


```{r, echo=FALSE}
p <- ggplot2::ggplot(acdt[36:55,], 
                     ggplot2::aes(factor(Time, levels = Time), 
                                  Temperature.F.))+
  ggplot2::geom_col(width = 0.2)+
  labs(title = "A traditional Line Chart Showing Accidents by Time", 
       x = "Time",
       y = "Temperature (°F)")
L2 <- p + ggplot2::theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
L2
```

When a small number of values are plotted, the line chart becomes much better
than its previous counterpart. There are, however, several limitations compared
to its clockplot counterpart.

a. It is hard to discern exactly when an event occurred (or recorded). What we
observe is just the trend of the temperature.  
b. It gives us an idea that recordings are placed at equal intervals. Actually,
the time gap between the first and the second recordings is around 8 hours, while
the next gap is worth 36 minutes only. Thus, any trend, either visual or analytical,
would be extremely faulty. 
c. Unlike the clockplot, we cannot single out the duration when there are
no, less, or large number of occurrences. Occurrences, as we have seen in
the previous clockplot, may be rare during night and early morning. We, however,
cannot extract that information from the line chart. 
d. When we plot other periods (day, month, year, and so on), they are not 
necessarily cyclic. We get a continuous flow of time, going ahead linearly: 
year 2015, year 2016, year 2017 etc; no year is repeated. Time data are, 
however, cyclic; after 24 hours, each hour starts repeating itself. A linear
line chart fails to reveal this cyclic behavior. As an example, the first and
second value might represent the end of a day (a cycle) and the start of a 
new day. A legend may help see this; clockplot still serves the purpose better.
In short, the line chart cannot reveal the gap between the last and first.

In the hospital data, for example, all the admissions take place between 2 Am
and 1 PM. A line chart will just plot the events between this period, without
revealing there were no admissions between 1 PM and 2 AM.
e. The line chart fails to show that the events are actually clustered around 
5 AM to 7 AM, as seen in the Figure \@ref{fig:acdt20}
f. If we have no values against times, for example, we only have timings
of events, such as when patients were admitted in a hospital, we cannot
present this on a line chart. To circumvent the problem, we may choose
an arbirary value against each time, but this does not allow us to focus
on the timings of the events.

```{r acdt20, fig.cap="CAP", echo=FALSE}
clock_chart(data = acdt[36:55,], time = Time)
```

```{r hospad, fig.cap="CAP", echo=FALSE}
hosp <- read.csv("https://raw.githubusercontent.com/mahmudstat/open-analysis/main/data/hospital.csv")
clock_chart_qlt(hosp[1:100,], Entry_Time, Financial_Class)+
  labs(color = "Patient Type")
```


Need cyclic
