---
title: "clockplot:A Novel Way to Visulaize Event Time Data"
author:
- name: Abdullah Al Mahmud
date: "`r Sys.Date()`"
output:
  html_document:
    fig_caption: true
  pdf_document: default
bibliography: references.bib 
---

# Structure

- Abstract
- Intro
- Methods
- Usage (functions, applications & interpretations, pattern detection)
- - Add faceted plots


# Abstract

  Visualization is a fundamental component of statistical analysis and prediction, enabling clear data description and informed model selection. While various methods exist for visualizing event data, there is a notable gap in representing events on a clock face. In this paper, we introduce clockplot, an R package for plotting timestamped events on a 24-hour circular clock face. This approach offers a powerful way to perceive the precise timing of events and facilitates intuitive comparisons with other events. The clockplot is particularly useful for analyzing daily patterns, event clustering, and temporal gaps. For event times, it is significantly more revealing than conventional visualizations like bar or pie charts. This package also generalizes the clockplot approach to create cyclic charts for other time frames, including weekly and monthly cycles. This functionality enables effective event planning and pattern analysis across multiple periods, providing a new and insightful tool for temporal data analysis.

# Introduction 

This paper revolves around plotting event time on a 24-hour clock, showing
events along their corresponding timings. The existing method for the similar
kind of data uses a line chart, also known as a time series graph or line 
graph sample ref [@R-base]. A line chart is drawn for illustrating a time 
series data, which is defined as a sequence of values at successive (usually 
equally spaced) points in time. A line chart is very broad in scope; it can
present time data sequenced by any unit of time, such as hour, day, week, month,
or year, and so on and so forth. 

Notationally, $Y$ is a time series data if it can be written as below:

$\displaystyle Y = (Yt: t \in T$

For most time series data encountered in practice, the Xi are equally spaced
through time. [@GMDA] 

The domains having the application of time series are numerous and include 
weather meteorology, economics, electroencephalography, control engineering, astronomy, communications engineering, geology, among others. 

One variation of a line chart is a run chart or run-sequence plot, which 
usually shows the performance of a process. As far as this paper is concerned,
they serve the same purpose and cannot serve what we are going to propose 
in this paper. 



## Clock Skeleton

## Plotting Times on the Clock

Let us explore US Accidents data. Although the main focus of these data 
seem to be on accident times, we can employ it to visualize the comparison 
between the existing line chart and our proposed clock chart. 

```{r}
library(clockplot)
library(ggplot2)
acdt <- read.csv("https://raw.githubusercontent.com/mahmudstat/open-analysis/main/data/usacc.csv")
acdt <- acdt[,c(9:12)]
head(acdt)
```

Let us plot temperature against time.

```{r lnct, fig.cap="CAP", echo=FALSE}
p <- ggplot2::ggplot(acdt, ggplot2::aes(Time, Temperature.F.))+
  ggplot2::geom_col(width = 0.2)+
  labs(title = "A traditional Line Chart Showing Accidents by Time", 
       y = "Temperature (°F)")
p + ggplot2::theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

We have no idea which temperature belongs to which time, although we have 
changed X- axis labels at right angles (i.e., vertically). 

Now let us take a look into a clock plot. 

```{r cl1f, ig.cap="CAP", echo=FALSE}
clock_chart(data = acdt, time = Time)
```

Although this is the simplest kind of clock chart that we propose, this still
makes a lot more sense. We clearly see a huge gap between around 9 PM to 
around 4:30 AM, during which period no reading of temperature was made (in
other example, this would mean no event occurred, which is actually relevant to
this data set as well, which records times of accidents. Why no accidents 
occurred during this period is subject to further discussion; perhaps, 
the traffic was less during this time.)

We can easily observe that most incidents occurred from 5 AM to 9 AM. Between 
other hours, roughly two accidents occurred. 


Let us try plotting 20 values only.


```{r lnt2, fig.cap="CAP", echo=FALSE}
p <- ggplot2::ggplot(acdt[36:55,], 
                     ggplot2::aes(factor(Time, levels = Time), 
                                  Temperature.F.))+
  ggplot2::geom_col(width = 0.2)+
  labs(title = "A traditional Line Chart Showing Accidents by Time", 
       x = "Time",
       y = "Temperature (°F)")
L2 <- p + ggplot2::theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
L2
```

When a small number of values are plotted, the line chart becomes much better
than its previous counterpart. There are, however, several limitations compared
to its clockplot counterpart.

a. It is hard to discern exactly when an event occurred (or recorded). What we
observe is just the trend of the temperature.  
b. It gives us an idea that recordings are placed at equal intervals. Actually,
the time gap between the first and the second recordings is around 8 hours, while
the next gap is worth 36 minutes only. Thus, any trend, either visual or analytical,
would be extremely faulty. 
c. Unlike the clockplot, we cannot single out the duration when there are
no, less, or large number of occurrences. Occurrences, as we have seen in
the previous clockplot, may be rare during night and early morning. We, however,
cannot extract that information from the line chart. 
d. When we plot other periods (day, month, year, and so on), they are not 
necessarily cyclic. We get a continuous flow of time, going ahead linearly: 
year 2015, year 2016, year 2017 etc; no year is repeated. Time data are, 
however, cyclic; after 24 hours, each hour starts repeating itself. A linear
line chart fails to reveal this cyclic behavior. As an example, the first and
second value might represent the end of a day (a cycle) and the start of a 
new day. A legend may help see this; clockplot still serves the purpose better.
In short, the line chart cannot reveal the gap between the last and first. 

  In the hospital data, for example, all the admissions take place between 2 Am
  and 1 PM. A line chart will just plot the events between this period, without
  revealing there were no admissions between 1 PM and 2 AM.
  
e. The line chart fails to show that the events are actually clustered around 
5 AM to 7 AM, as seen in the Figure \@ref{fig:acdt20}.
f. If we have no values against times, for example, we only have timings
of events, such as when patients were admitted in a hospital, we cannot
present this on a line chart. To circumvent the problem, we may choose
an arbirary value against each time, but this does not allow us to focus
on the timings of the events.

```{r acdt20, fig.cap="CAP", echo=FALSE}
clock_chart(data = acdt[36:55,], time = Time)
```

```{r hospad, fig.cap="CAP", echo=FALSE}
hosp <- read.csv("https://raw.githubusercontent.com/mahmudstat/open-analysis/main/data/hospital.csv")
clock_chart_qlt(hosp[1:100,], Entry_Time, Financial_Class)+
  labs(color = "Patient Type")
```


Need cyclic

# Methodology

Broadly speaking, the construction of the clockcharts have two steps. 

i. Construction of the 24-hour clock
ii. Plotting event times on the clock

Each steps has two substeps:

a. The mathematical idea
b. The programming

### The Clock Skeleton

First we need to construct the skeleton, the clock on which event times are 
plotted. For this, we need a circle, on which we have to mark 24 
dots to denote 24 hours, starting from 00:00 (12 AM) to 23:00 (11 PM). 

It is enough to draw a unit circle, although a circle of any radius would do. 
In `R`, it is easiest to draw a circle using the complex equation. For a unit
circle, the equation is $e^{i\theta}$, where $e$ is Euler's number, an 
irrational constant having the approximate value of 2.71828; $i$ stands for 
imaginary unit, equal to $\sqrt{-1}$; and $\theta$ refers to the trigonometric
angle in radian. To make a complete circle, we need to allow $\theta$ to
take values from 0 to 360 degrees ($2\pi$ in radian). 

Programmatically, we need to create some angles ranging from 0 to $2\pi$ and 
then connect them to make a complete circle. Let us create 100 angles and use
them in the circle equation ($e^{i\theta}$). The trigonometric form of the 
equation is $sin^2 \theta + cos^2 \theta = 1$, while the algebraic form is
$x^2+y^2 = 1$. The complex form, however, is more convenient. 

```{r, comc, fig.height=4, fig.width=4, fig.align='center', fig.cap="The Unit Circle"}
k <- 100
timepoint <- exp(1i * 2 * pi * (k:1) / k)
plot(timepoint, pch = 19, type = "b")
```

Here `R` just takes the complex numbers and plots the values of the real part
on the X axis and the corresponding values of the imaginary part on the Y axis. 

Now we have to mark the hours on the circle to make it look like a clock. We 
have to show 24 hours marks as well some marks for minute marks, corresponding 
to the fraction of the hour. This clock is slightly different from a traditional
clock not only in that it has 24 hours instead of 12, but also that it has 
only one hand, the hour hand, with minute part added to the hour mark as the 
fraction; the second part, having minor contribution, is ignored. 

First, let us see the output with 24 hour marks. 

```{r circle, fig.height=4, fig.width=4, fig.align='center'}
k <- 24
hour <- exp(1i * 2 * pi * (k:1) / k)
plot(hour, pch = 19)
```

At this time, we move to the `R` package `ggplot2` to produce better graphics. 
We need to create an `R` data frame for plotting ggplots. We also mark some 
points for minutes. 

First, let us create the required data frame

```{r, fig.height=4, fig.width=4, fig.align='center'}
k <- 24 # Hours
subk <- 24*4 # Fraction of hours
times <- exp(1i * 2 * pi * (k:1) / k)
subtimes <- data.frame(SubT = exp(1i * 2 * pi * (subk:1) / subk)) 
ampm = c(rep(" AM",6), rep(" PM",12), rep(" AM",6))
library(tibble)
dfclock <- tibble(time = times,
                  hour = c(6:12, 1:12, 1:5), # May not be needed
                  label = paste0(c(6:12, 1:5), ampm))
```

**Explanation**

`k = 24` indicates there are 24 hours. Then in each hour is divided into 4 parts
, corresponding to the elapse of successive 15 minutes. Thus, a small point on 
\@ref(fig:clock24) in the middle of two larger point (hour) means half an hour
has elapsed. For instance, the the point in the middle of 7 PM and 8 PM is 7:30 
PM. 

Now the code `exp(1i * 2 * pi * (k:1) / k)` gives us all the angles 
corresponding to 24 hours on the clock. Here, we have to note that the angles
in trigonometry starts from $0^o$ (X-axis) and rotates counter-clockwise, the 
direction opposite to the rotation of clock hands (the significance of the
difference will be seen shortly). But the clock moves clockwise. 
At 0 degree, we have 3 in a 12-hour clock and 6 in 24-hours clock.

Similarly, the code `SubT = exp(1i * 2 * pi * (subk:1) / subk)` gives us the
angles corresponding to the fraction of an hour. As we have already mentioned, 
the clock hand rotates clock-wise, starting from the $0^o$ angle, so we name 
the labels 6 PM to 11 PM, then we have 12 PM to 11 PM, followed by 12 AM to 
5 AM. 

The code `c(rep(" AM",6), rep(" PM",12), rep(" AM",6))` ensures this sequence. 

We then place the data into two data frames `subtimes` and `dfclock`, since
they are unequal dimensions. 

```{r skeleton, fig.height=4, fig.width=4, fig.align='center'}
library(ggplot2) # Load the package ggplot2
p1 <- dfclock %>% ggplot()+
  geom_path(data = subtimes, aes(Re(SubT), Im(SubT)))+
  # Connect Last two missing points
  geom_line(data = subtimes %>% dplyr::slice(-c(2:95)), aes(Re(SubT), Im(SubT)))+
  theme(aspect.ratio = 1)+
  geom_text(data = dfclock, aes(Re(time)*1.1, Im(time)*1.1, label = label))+
  geom_point(data = subtimes, 
             aes(Re(SubT), Im(SubT)), shape = 19, color = "black", size = 0.6)+
  geom_point(aes(Re(time), Im(time)), color = "black", size = 1.8)
p1
```

Next, we plot the lines using the `ggplot2` package. The functions `geom_path()`
does the job, while `geom_line()` is used to fill the gap between the first and 
the last point. `aspect.ratio = 1` is set to ensure the chart size do not 
depend on resizing of `Rstudio` Viewer size.

The points corresponding to the hours are labelled AM or PM using the 
`geom_text()` function. We do not need the axis labels and ticks. Let us 
remove them:

```{r skeleton_clean, fig.height=4, fig.width=4, fig.align='center'}
p1 + theme(axis.text.x=element_blank(), 
        axis.ticks.x=element_blank(), 
        axis.text.y=element_blank(),  
        axis.ticks.y=element_blank())+
  labs(x = "", y = "")
```

This completes our first part of the plot; we get the skeleton of the clock. 

The second and final part of the task involves plotting event times on the
appropriate points. Now, this task has several variations. In the simplest 
form, we just have a chart, so called `clock_chart`, which plots the event
times on the 24-hour clock. An optional argument `Col` can be used to change
the color of the clock hands. The other variations will be discussed later, when
we complete the discussion on how the event times are prepared for plotting.

#### Preparation of Event Times Data

We have data in the format `HH:MM:SS`. We need to convert this data to `(x,y)`
coordinates. In the unit circle, for example, the 12 AM should be plotted on
the Y-axis (x = 0, y = 1) and 6 AM on the X axis (x = 1, y = 0), although, to
avoid plotting on the circumference, we decrease the length a little (more on
this later). 

To convert the `HH:MM:SS` to coordinates of the unit circle, we first separate 
the parts of time into `HH`, `MM`, and `SS`. We accomplish that using the 
`dplyr` and `tidyr` package. 

```{r}
# Data to check on
chkdf <- data.frame(time = c("06:00:00", "12:00:00", "17:30:00", "00:05:25"),
                    value = sample(10,4))
library(dplyr)
chkdf %>% tidyr::separate_wider_delim(cols = time, 
                         names = c("hour", "minute", "second"), 
                         cols_remove = FALSE,
                         delim = ":")
```

This gives us `hour`, `minute`, and `second` as character (`chr`) vectors, but 
we need them as numeric vectors. The conversion is made with `as.numeric` 
function. 

Here the challenge now is to combine the hour and the minute into one singe
time, which we would later convert to coordinate. We intentionally ignore the
second (`SS`), since it has too negligible contribution on the plot.

Here a short algorithm is to be used:

Note here that 30 minutes is equivalent to half an hour, so `1:30:00` corresponds
to 1.5 hour. We need to recall that time values range from 0 to 60, while 
our decimal number system gives us the suitable range 1 to 100. In our number
system, reaching 100 indicates we have completed some work, while clockworks 
say reaching 60 completes one cycle of the task. 

The hour part does not need conversion. 23 is 23 and 22 is 22. 

Thus, 15 minutes on the clock says $\frac{15}{60} = \frac14$ th or $25\%$ work, 
30 means $\frac{30}{60} = \frac12 = 0.5$ or $50\%$ work, and 50 means 
$\frac{50}{60}$ or $83\%$ work have been completed. 

The task is further complicated due to the clock and angles rotating in
opposite directions. 

////THIS CONVERSION IS A BUG////
ifelse(.data$minute<10, .data$minute * 5/30, .data$minute * 5/300)
////MAY DEPEND ON OLD TIME FORMAT HHMM hrs/////




`r if (knitr::is_html_output()) '# References {-}'`